---
output:
  md_document:
    variant: markdown_github
---

```{r options, echo=FALSE}
knitr::opts_chunk$set(fig.path = "", message = FALSE, warning = FALSE)
```

### Introduction

The `eurostatR` package package provides an `R` interface to the [Eurostat API](https://ec.europa.eu/eurostat).

To install the development version:

```{r install, eval=FALSE}
library(devtools)
install_github("expersso/eurostatR")
```

### Example usage

Start by retrieving a dataframe with all available data sets:

```{r}
library(eurostatR)

flows <- get_dataflows()
head(flows)
```

Now, let's say we're interested in the share of employees with a doctorate,
broken down by occupation:

```{r}
flows[grep("doctorate", flows$description),]
id <- "cdh_e_occ1"
```

Before getting the data, we need to know the available dimensions for this data
set:

```{r}
dims <- get_dimensions(id)
str(dims)
```

The last value of the returned list (`dims$key`) gives the general form of the
key that we need to request. By inspecting the dimensions in the list, we find
that `FREQ` should be `A` (annual), `Y_GRAD` should be `TOTAL`, and `UNIT`
should be `PC` (percent). The dimension `ISCO88` gives the different
occupations, and since we want all of these, we leave that dimension empty.
Finally we want to look at date for Germany and the US (`DE` and `US`), which we
specify with a `+`.

```{r}
key <- "cdh_e_occ1.A.TOTAL.PC..DE+US"
doctorates <- get_data(key)
head(doctorates)
```

We can now plot these data as a parallel coordinates plot:

```{r plot, fig.width = 6, fig.height = 4}
library(dplyr)
library(ggplot2)

doctorates <- left_join(doctorates, dims$dimensions$ISCO88, 
                        by = c("isco88" = "id"))

# Factor order for occupation variable, in descending order
lvls <- doctorates %>% 
  filter(geo == "US") %>% 
  arrange(obsvalue) %>% 
  .$description

doctorates %>% 
  filter(obstime == 2006) %>% 
  filter(!is.na(obsvalue)) %>% 
  filter(description != "Professionals") %>% 
  ggplot(aes(x = factor(description, lvls), 
             y = obsvalue, fill = geo, color = geo, group = geo)) +
  geom_line() +
  geom_point() +
  coord_flip() +
  theme_light(8) +
  labs(y = "% share of employees", x = NULL,
       title = "Employed doctorate holders by occupation")
```

The chart tells us e.g. that doctorates are much more prevalent among teaching
professionals in the US than they are in Germany.

### Disclaimer

This package is in no way officially related to, or endorsed by, the Eurostat.